---
title: 'Acoustic features analysis'
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(magrittr)
library(corrplot)
library(broom)
```

This script explores the acoustic features in the expanded corpus, specifically:

- correlations between features
- correlations between individual feature differences and similarity between originals and covers
- PCA to find low-dimensional representation that best captures variability, both on absolute features and scaled feature differences


First, load the features and feature differences (unscaled).
```{r}
features <- read_csv('../../data/processed/expanded/joined_with_features/all_features.csv')
feature_diffs <- read_csv('../../data/processed/expanded/joined_with_features/similarity_and_feature_differences.csv')
```

Select only columns with acoustic features.
```{r}
acoustic_feat <- features %>% select(1:21)
acoustic_diffs <- feature_diffs %>% select(1:24)
```

\newpage
# Correlation between features
## Features
Compute the correlation matrix.
```{r}
corr_matrix <- cor(acoustic_feat[-1]) # leave out song_id
```

```{r}
corrplot(corr_matrix, type = 'upper', # upper triangle
         order = 'hclust', # automatically orders with hierarchical clustering
         tl.col = 'black', tl.srt = 45) # specify text color and rotation
```
\newpage
Without MFCCs.
```{r}
corr_matrix_subset <- cor(select(acoustic_feat, -c(song_id, contains("mfcc"))))
# leave out song_id, MFCCs
```

```{r}
corrplot(corr_matrix_subset, type = 'upper', # upper triangle
         method = 'number', # see actual correlation values
         order = 'hclust', # automatically orders with hierarchical clustering
         tl.col = 'black', tl.srt = 45) # specify text color and rotation
```
\newpage
Now just look at MFCCs.
```{r}
corr_matrix_MFCCs <- cor(select(acoustic_feat, contains("mfcc")))
# keep only MFCCs
```

```{r}
corrplot(corr_matrix_MFCCs, type = 'upper', # upper triangle
         #method = 'number', # see actual correlation values
         order = 'hclust', # automatically orders with hierarchical clustering
         tl.col = 'black', tl.srt = 45) # specify text color and rotation
```

\newpage
## Feature differences
Are there correlations between feature differences?
```{r}
corr_matrix_diffs <- cor(acoustic_diffs[-seq(1,3)]) # leave out song_id, similarity
```

```{r}
corrplot(corr_matrix_diffs, type = 'upper', # upper triangle
         order = 'hclust', # automatically orders with hierarchical clustering
         tl.col = 'black', tl.srt = 45) # specify text color and rotation
```

Yes, there seem to be similar systematic differences.

\newpage
```{r}
corr_matrix_subset_diffs <- cor(select(acoustic_diffs, -c(seq(1,3), contains("mfcc"))))
```

```{r}
corrplot(corr_matrix_subset_diffs, type = 'upper', # upper triangle
         method = 'number',
         order = 'hclust', # automatically orders with hierarchical clustering
         tl.col = 'black', tl.srt = 45) # specify text color and rotation
```

\newpage
```{r}
corr_matrix_mfcc_diffs <- cor(select(acoustic_diffs, contains("mfcc")))
```

```{r}
corrplot(corr_matrix_mfcc_diffs, type = 'upper', # upper triangle
         #method = 'number',
         order = 'hclust', # automatically orders with hierarchical clustering
         tl.col = 'black', tl.srt = 45) # specify text color and rotation
```

\newpage
# Correlations between individual feature differences and similarity
How well do individual features predict similarity?
Take these results with a grain of salt because we're running A LOT of models.
```{r}
summary(lm(mean_sim ~ rms_difference, acoustic_diffs))
summary(lm(mean_sim ~ spectral_bandwidth_difference, acoustic_diffs))
summary(lm(mean_sim ~ spectral_centroid_difference, acoustic_diffs))
summary(lm(mean_sim ~ spectral_contrast_difference, acoustic_diffs))
summary(lm(mean_sim ~ spectral_flatness_difference, acoustic_diffs))
summary(lm(mean_sim ~ spectral_rolloff_difference, acoustic_diffs))
summary(lm(mean_sim ~ zero_crossing_rate_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfccs_offset_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc1_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc2_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc3_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc4_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc5_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc6_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc7_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc8_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc9_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc10_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc11_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfcc12_difference, acoustic_diffs))
summary(lm(mean_sim ~ mfccs_dist, acoustic_diffs))
```

Well, none of the acoustic features individually predict similarity except MFCCs distance. This is not super surprising.

\newpage
# Predicting similarity with all acoustic features

Run the full and subset models just to get a sense...
```{r}
summary(lm(mean_sim ~ rms_difference + spectral_bandwidth_difference + 
             spectral_centroid_difference + spectral_contrast_difference +
             spectral_flatness_difference + spectral_rolloff_difference +
             zero_crossing_rate_difference, 
           acoustic_diffs))
```

```{r}
summary(lm(mean_sim ~ mfccs_offset_difference + mfcc1_difference + mfcc2_difference + mfcc3_difference + mfcc4_difference + mfcc5_difference + mfcc6_difference + mfcc7_difference + mfcc8_difference + mfcc9_difference + mfcc10_difference + mfcc11_difference + mfcc12_difference, 
           acoustic_diffs))
```

```{r}
summary(lm(mean_sim ~ rms_difference + spectral_bandwidth_difference + 
             spectral_centroid_difference + spectral_contrast_difference +
             spectral_flatness_difference + spectral_rolloff_difference +
             zero_crossing_rate_difference + mfccs_offset_difference + mfcc1_difference + mfcc2_difference + mfcc3_difference + mfcc4_difference + mfcc5_difference + mfcc6_difference + mfcc7_difference + mfcc8_difference + mfcc9_difference + mfcc10_difference + mfcc11_difference + mfcc12_difference, 
           acoustic_diffs))
```

Cool, so NOTHING is significant.

\newpage
# Principal components
(Actual PCA performed in `acoustic_PCA.ipynb`, just analyzed here...)

Load the PCs for analysis.
```{r}
acoustic_PCs <- read_csv('../../data/processed/expanded/joined_with_features/acoustic_PCs.csv')
```

The first 6 PCs explains about 80% of the variance in the acoustic features.

Pivot the data wider and calculate differences in the first 6 PCs.
```{r}
pcs_diff <- acoustic_PCs %>%
  select(seq(1:7)) %>%
  # prepare the data to pivot wider
  separate_wider_delim(song_id, delim = '_', names = c('title', 'artist', 'orig_or_cover')) %>%
  mutate(song_id = paste(title, artist, sep = '_')) %>%
  relocate(song_id) %>%
  mutate(title = NULL, artist = NULL) %>%
  # pivot longer first, so each row is one PC for one clip
  pivot_longer(cols = -c(song_id, orig_or_cover), names_to = 'PC', values_to = 'value') %>%
  # pivot wider to group by song and PC
  pivot_wider(id_cols = c(song_id, PC), names_from = orig_or_cover, values_from = value) %>%
  # calculate the difference between each PC for each song
  mutate(difference = Original5S - Cover5S) %>%
  # drop original and cover (saved elsewhere)
  select(-c(Original5S, Cover5S)) %>%
  # add 1 to the PCs column to make it 1-6 (rather than 0-5)
  mutate(PC = as_factor(as.numeric(PC) + 1)) %>%
  # pivot wider to get back to each song has one row
  pivot_wider(id_cols = song_id, names_from = PC, values_from = difference,
              names_glue = "PC{PC}_difference")
```

Save this dataframe.
```{r}
#write_csv(pcs_diff, '../../data/processed/expanded/joined_with_features/acoustic_PC_diffs.csv')
```

The PCs themselves should be uncorrelated (which they are).

```{r}
corr_matrix_PCs <- cor(acoustic_PCs[-1])
corrplot(corr_matrix_PCs, type = 'upper', 
         tl.col = 'black', tl.srt = 45) 
```

\newpage
But the differences might be correlated?

```{r}
corr_matrix_PCs_diff <- cor(pcs_diff[-1])
corrplot(corr_matrix_PCs_diff, type = 'upper', 
         tl.col = 'black', tl.srt = 45) 
```

\newpage
Try to predict similarity with all 6 PCs individually.
```{r}
pcs_diff_sim <- acoustic_diffs %>%
  select(song_id, mean_sim) %>%
  left_join(., pcs_diff, by = join_by(song_id))
```

```{r}
summary(lm(mean_sim ~ abs(PC1_difference), pcs_diff_sim))
summary(lm(mean_sim ~ abs(PC2_difference), pcs_diff_sim))
summary(lm(mean_sim ~ abs(PC3_difference), pcs_diff_sim))
summary(lm(mean_sim ~ abs(PC4_difference), pcs_diff_sim))
summary(lm(mean_sim ~ abs(PC5_difference), pcs_diff_sim))
summary(lm(mean_sim ~ abs(PC6_difference), pcs_diff_sim))
```

\newpage
And together.
```{r}
summary(lm(mean_sim ~ abs(PC1_difference) + abs(PC2_difference) +
             abs(PC3_difference) + abs(PC4_difference) +
             abs(PC5_difference) + abs(PC6_difference), pcs_diff_sim))
```

\newpage
Visualize just the first PC.

```{r}
ggplot(pcs_diff_sim, aes(x = abs(PC1_difference), y = mean_sim)) +
  geom_point()
```


\newpage
# PCA on scaled feature differences
Load the PCs of the scaled feature differences.
```{r}
diffs_PCs <- read_csv('../../data/processed/expanded/joined_with_features/acoustic_diffs_PCs_noMFCCs.csv')
```

The PCs themselves should be uncorrelated (which they are).

```{r}
corr_matrix_PCs <- cor(diffs_PCs[-1])
corrplot(corr_matrix_PCs, type = 'upper', 
         tl.col = 'black', tl.srt = 45) 
```

Load mean similarity ratings.
```{r}
sim_ratings <- read_csv('../../data/processed/expanded/joined_with_features/similarity_and_feature_scaled_differences.csv')
sim_ratings %<>% select(c(song_id, mean_sim))
diffs_PCs %<>%
  left_join(sim_ratings, ., by=join_by(song_id))
```

\newpage
Run models for each PC individually (using `broom`).
```{r}
# code adapted from this post: https://stackoverflow.com/questions/54907726/running-multiple-linear-regressions-across-several-columns-of-a-data-frame-in-r
individual_models <- diffs_PCs %>%
  select(-song_id) %>%
  pivot_longer(cols = -mean_sim,
               names_to = 'column', values_to = 'value') %>%
  nest(data = -column) %>% # this gives us one model per PC
  # each column gets a tibble with mean_sim and value as columns
  mutate(model = map(data, ~lm(mean_sim ~ value, data = .))) %>% # run the models
  mutate(tidy_model = map(model, tidy)) %>% # tidy the models
  select(-data, -model) %>% # remove untidy models
  unnest(cols = c(tidy_model))
print(individual_models)
```

PCs 1 has a marginally significant p-values in their individual models (between 0.05 and 0.10). However, we're running 7 separate models so that p-value probably would not survive multiple comparisons corrections.


Run one model with all 7 PCs.
```{r}
# help from here: https://stackoverflow.com/questions/67665340/incorporate-all-columns-of-a-dataframe-into-one-regression
model20 <- diffs_PCs %>%
  select(-song_id) %>%
  lm(mean_sim ~ ., data = .) # first . means all other columns
summary(model20)
```

Now the first and third PC are marginally significant. The adjusted R^2 (0.057) is quite a bit lower than R^2 (0.153).

