---
title: "Variance partitioning"
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
library(tidyverse)
library(magrittr)
library(vegan) # to perform variance partitioning
library(eulerr) # to make area-proportional Venn diagrams
select <- dplyr::select
```

This notebook uses the musical and emotional features and performs variance partitioning to understand how these sets of features work together to explain similarity.

Load the data.
```{r}
feature_diffs_scaled <- read_csv('../../data/processed/expanded/joined_with_features/similarity_and_feature_scaled_differences.csv')
```

Keep the musical and emotional features.
```{r}
data <- feature_diffs_scaled %>% select(1:2, 23:25, 28:30)
```

With help from: https://www.davidzeleny.net/anadat-r/doku.php/en:varpart

Venn diagram parts:

- a: variance explained by musical features only
- b: variance explained by musical and emotional features together
- c: variance explained by emotional features only
- d: unexplained variance

For the musical features, we'll allow key, voice type, and MFCCs to interact because that was the best musical features model (see `musical_features.Rmd`).

\newpage
# Calculate manually

[a+b+c]
```{r}
rda.all <- rda(data$mean_sim ~ tempo_difference + key_difference*voice_type_difference*mfccs_dist +
                 mean_valence_difference + mean_arousal_difference,
               data)
summary(rda.all)
```

Is this consistent with just a simple linear model?
```{r, include = FALSE}
summary(lm(mean_sim ~ tempo_difference + key_difference*voice_type_difference*mfccs_dist +
             mean_valence_difference + mean_arousal_difference,
           data))
```

Yes, the explained variance is 0.5821, which is similar to the "constrained" variance in the RDA model. For RDA, it seems like "unconstrained" variance is [d].


Model for musical features only [a+b]
```{r}
rda.mus <- rda(data$mean_sim ~ tempo_difference + key_difference*voice_type_difference*mfccs_dist,
               data)
summary(rda.mus)
```

Model for emotional features only [b+c]
```{r}
rda.emot <- rda(data$mean_sim ~ mean_valence_difference + mean_arousal_difference,
               data)
summary(rda.emot)
```

Extract the R^2 values.

```{r}
abc <- RsquareAdj(rda.all)$adj.r.squared
ab <- RsquareAdj(rda.mus)$adj.r.squared
bc <- RsquareAdj(rda.emot)$adj.r.squared
```

Subtract to find individual fractions.

```{r}
# shared variance
b <- ab + bc - abc
# variance explained by musical features only
a <- ab - b
# variance explained by emotional features only
c <- bc - b
```

- variance explained by musical features only (a): 20.58% 
- shared variance explained (b): 29.74%
- variance explained by emotional features only (c): 7.89%
- unexplained variance (d): 41.79%


\newpage
# Musical features

- a: variance explained by tempo only
- b: shared variance explained by the musical features
- c: variance explained by key, voice type, and MFCCs only
- d: unexplained variance

[a+b+c]
```{r}
rda.mus <- rda(data$mean_sim ~ tempo_difference + key_difference*voice_type_difference*mfccs_dist,
               data)
summary(rda.mus)
```

[a+b]
```{r}
rda.tempo <- rda(data$mean_sim ~ tempo_difference,
               data)
summary(rda.tempo)
```

[b+c]
```{r}
rda.other <- rda(data$mean_sim ~ key_difference*voice_type_difference*mfccs_dist,
               data)
summary(rda.other)
```

Extract the R^2 values.

```{r}
abc_m <- RsquareAdj(rda.mus)$adj.r.squared
ab_m <- RsquareAdj(rda.tempo)$adj.r.squared
bc_m <- RsquareAdj(rda.other)$adj.r.squared
```

Subtract to find individual fractions.

```{r}
# shared variance
b_m <- ab_m + bc_m - abc_m
# variance explained by tempo only
a_m <- ab_m - b_m
# variance explained by key, voice type, and timbre only
c_m <- bc_m - b_m
```


- variance explained by tempo only (a): 26.28% 
- shared variance (b): 6.61%
- variance explained by key, voice type, and timbre only (c): 17.42%
- unexplained variance (d): 49.69%

\newpage
# Emotional features

- a: variance explained by valence only
- b: shared variance explained by the emotional features
- c: variance explained by arousal only
- d: unexplained variance

[a+b+c]
```{r}
rda.emot <- rda(data$mean_sim ~ mean_valence_difference + mean_arousal_difference,
               data)
summary(rda.emot)
```

[a+b]
```{r}
rda.val <- rda(data$mean_sim ~ mean_valence_difference,
               data)
summary(rda.val)
```

[b+c]
```{r}
rda.arous <- rda(data$mean_sim ~ mean_arousal_difference,
               data)
summary(rda.arous)
```

Extract the R^2 values.

```{r}
abc_e <- RsquareAdj(rda.emot)$adj.r.squared
ab_e <- RsquareAdj(rda.val)$adj.r.squared
bc_e <- RsquareAdj(rda.arous)$adj.r.squared
```

Subtract to find individual fractions.

```{r}
# shared variance
b_e <- ab_e + bc_e - abc_e
# variance explained by valence only
a_e <- ab_e - b_e
# variance explained by arousal only
c_e <- bc_e - b_e
```


- variance explained by valence only (a): 18.51%
- shared variance by the emotional features (b): 10.92%
- variance explained by arousal only (c): 8.20%
- unexplained variance (d): 62.37%

\newpage
# Using `varpart`

Musical and emotional features as a set
```{r}
varp_2sets <- varpart(data$mean_sim, 
                      ~ tempo_difference + key_difference*voice_type_difference*mfccs_dist,
                      ~ mean_valence_difference + mean_arousal_difference, 
                      data=data)
varp_2sets
```

Plot the results with a Venn diagram

```{r}
plot(varp_2sets, digits = 2, Xnames = c('musical\nfeatures', 'emotional\nfeatures'))
```

Plot an area-proportional Venn diagram

```{r}
these_fractions <- varp_2sets$part$indfract$Adj.R.squared
combo <- c(A=these_fractions[1], B=these_fractions[2], "A&B"=these_fractions[3])
percents <- round(combo * 100, digits = 2)
percents <- paste(percents, "%", sep = '')

this_fit <- euler(combo, shape = "ellipse")
plot(this_fit, fills = c('#1F78B4','#F14D0E'), quantities = percents,
     labels = c('musical features', 'emotional features'))
```


Separating into three sets (tempo, all other musical features, emotional features)

```{r}
varp_3sets <- varpart(data$mean_sim, 
                      ~ tempo_difference,
                      ~ key_difference*voice_type_difference*mfccs_dist,
                      ~ mean_valence_difference + mean_arousal_difference, 
                      data=data)
varp_3sets
```

```{r}
plot(varp_3sets, digits = 2, Xnames = c('tempo', 'other musical\nfeatures', 'emotional\nfeatures'))
```

```{r}
these_fractions <- varp_3sets$part$indfract$Adj.R.square
these_fractions <- ifelse(these_fractions < 0, 0, these_fractions)

combo <- c(A=these_fractions[1], B=these_fractions[2], C=these_fractions[3],
           "A&B"=these_fractions[4], 
           "A&C"=these_fractions[6], # order matters apparently
           "B&C"=these_fractions[5], 
           "A&B&C"=these_fractions[7])
percents <- round(combo * 100, digits = 2)
percents <- paste(percents, "%", sep = '')

this_fit <- euler(combo, shape = "ellipse")
p3 <- plot(this_fit, fills = c('#825AA9', '#298C70','#F14D0E'), 
          alpha = 0.8, quantities = percents,
          labels = c('tempo', 'key,\nvoice type,\ntimbre', 'valence & arousal'))
p3
#ggsave('../figures/venn_three-way.png', p3)
```

\newpage

Venn diagrams for musical features only and emotional features only.

Musical features only
```{r}
varp_mus <- varpart(data$mean_sim, 
                    ~ tempo_difference,
                    ~ key_difference*voice_type_difference*mfccs_dist,
                    data=data)
varp_mus
```

Plot the results with a Venn diagram

```{r}
plot(varp_mus, digits = 2, Xnames = c('tempo', 'key, voice type, timbre'))
```

Plot an area-proportional Venn diagram

```{r}
these_fractions <- varp_mus$part$indfract$Adj.R.squared
combo <- c(A=these_fractions[1], B=these_fractions[2], "A&B"=these_fractions[3])
percents <- round(combo * 100, digits = 2)
percents <- paste(percents, "%", sep = '')

this_fit <- euler(combo, shape = "ellipse")
p.mus <- plot(this_fit, fills = c('#825AA9','#298C70'), 
              quantities = percents, alpha = 0.8,
              labels = c('tempo', 'key, voice type, timbre'))
p.mus
#ggsave('../figures/venn_musical.png', p.mus)
```


Emotional features only
```{r}
varp_emot <- varpart(data$mean_sim, 
                    ~ mean_valence_difference,
                    ~ mean_arousal_difference,
                    data=data)
varp_emot
```

Plot the results with a Venn diagram

```{r}
plot(varp_emot, digits = 2, Xnames = c('valence', 'arousal'))
```

Plot an area-proportional Venn diagram

```{r}
these_fractions <- varp_emot$part$indfract$Adj.R.squared
combo <- c(A=these_fractions[1], B=these_fractions[2], "A&B"=these_fractions[3])
percents <- round(combo * 100, digits = 2)
percents <- paste(percents, "%", sep = '')

this_fit <- euler(combo, shape = "ellipse")
p.emot <- plot(this_fit, fills = c('#FF7F00','#E31A1C'), 
              quantities = percents, alpha = 0.8,
              labels = c('valence', 'arousal'))
p.emot
#ggsave('../figures/venn_emotional.png', p.emot)
```

\newpage
Can we do this with valence and arousal separately in a four-way model?

```{r}
varp_4 <- varpart(data$mean_sim, 
                  ~ tempo_difference,
                  ~ key_difference*voice_type_difference*mfccs_dist,
                  ~ mean_valence_difference,
                  ~ mean_arousal_difference, 
                  data=data)
varp_4
```

Plot the results with a Venn diagram

```{r}
plot(varp_4, digits = 2, Xnames = c('tempo', 'other musical features',
                                       'valence', 'arousal'))
```

Try to plot an area-proportional Venn diagram

```{r}
these_fractions <- varp_4$part$indfract$Adj.R.square
these_fractions <- ifelse(these_fractions < 0, 0, these_fractions)

combo <- c(t=these_fractions[1], omf=these_fractions[2], 
           v=these_fractions[3], a=these_fractions[4],
           "t&omf"=these_fractions[5], 
           "t&v"=these_fractions[7],
           "t&a"=these_fractions[8],
           "omf&v"=these_fractions[6],
           "omf&a"=these_fractions[9], 
           "v&a"=these_fractions[10],
           "t&omf&a"=these_fractions[11], 
           "t&omf&v"=these_fractions[12],
           "t&v&a"=these_fractions[14],
           "omf&v&a"=these_fractions[13], 
           "t&omf&v&a"=these_fractions[15])
percents <- round(combo * 100, digits = 2)
percents <- paste(percents, "%", sep = '')

this_fit <- euler(combo, shape = "ellipse")
p4 <- plot(this_fit, fills = c('#825AA9','#298C70','#FF7F00','#E31A1C'), 
              quantities = percents, alpha = 0.8,
              labels = c('tempo', 'key,\nvoice type,\ntimbre', 'valence', 'arousal'))
p4
#ggsave('../figures/venn_four-way.png', p4)
```