---
title: "Emotional features to memory"
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

This notebook looks at the emotional ratings of memories as a function of the emotional valence and arousal of the songs they are evoked by.

```{r}
library(tidyverse)
library(magrittr)
library(lme4)
library(lmerTest)
```

Load the data.
```{r}
data <- read_csv('../../data/processed/dense/joined_with_features/memory_all.csv')
```

Make the memory ratings factors.
```{r}
common_rating <- c('not at all', 'a little', 'somewhat', 'very', 'extremely')
data_factor <- data %>%
  mutate(emot_content = factor(emot_content,
                               levels = c('very negative','somewhat negative',
                                          'neutral','somewhat positive',
                                          'very positive'),
                               ordered = TRUE),
         emot_intensity = factor(emot_intensity, 
                                 levels = common_rating,
                                 ordered = TRUE),
         energetic = factor(energetic,
                            levels = common_rating,
                            ordered = TRUE)
         )

```


Predict emotional valence of the memories using emotional valence and arousal of the clips.

```{r}
summary(lmer(as.numeric(emot_content) ~ mean_valence + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(mean_valence, as.numeric(emot_content))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```

```{r}
summary(lmer(as.numeric(emot_content) ~ mean_arousal + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(mean_arousal, as.numeric(emot_content))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```

Predict emotional intensity of the memories using emotional valence and arousal of the clips.

```{r}
summary(lmer(as.numeric(emot_intensity) ~ mean_valence + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(mean_valence, as.numeric(emot_intensity))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```

```{r}
summary(lmer(as.numeric(emot_intensity) ~ mean_arousal + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(mean_arousal, as.numeric(emot_intensity))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```

Predict energy of the memories using emotional valence and arousal of the clips.

```{r}
summary(lmer(as.numeric(energetic) ~ mean_valence + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(mean_valence, as.numeric(energetic))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```

```{r}
summary(lmer(as.numeric(energetic) ~ mean_arousal + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(mean_arousal, as.numeric(energetic))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```

\newpage

Valence and arousal together

```{r}
summary(lmer(as.numeric(emot_content) ~ mean_valence*mean_arousal + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
summary(lmer(as.numeric(emot_intensity) ~ mean_valence*mean_arousal + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
summary(lmer(as.numeric(energetic) ~ mean_valence*mean_arousal + (1|internal_id) + (1|song_id), data_factor))
```



\newpage
Tempo?

```{r}
summary(lmer(as.numeric(emot_content) ~ tempo + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(tempo, as.numeric(emot_content))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```

```{r}
summary(lmer(as.numeric(emot_intensity) ~ tempo + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(tempo, as.numeric(emot_intensity))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```

```{r}
summary(lmer(as.numeric(energetic) ~ tempo + (1|internal_id) + (1|song_id), data_factor))
```

```{r}
ggplot(data_factor, aes(tempo, as.numeric(energetic))) +
  geom_jitter(width = 0) +
  geom_smooth(method = 'lm')
```
