---
title: "Memory likelihood"
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

The notebook looks at the likelihood of a cover memory, given that a memory was evoked by the original. We try to predict that given perceptual similarity and musical/emotional features.

```{r}
library(tidyverse)
library(magrittr)
library(lme4)
library(lmerTest)
```

Load the data.
```{r}
data <- read_csv('../../data/processed/dense/joined_with_features/memory_original-w_or_wo_cover_memories.csv')
```

Load just the features and similarity for song-wise analysis.
```{r}
by_song <- read_csv('../../data/processed/dense/joined_with_features/similarity_and_scaled_feature_differences.csv')
```


\newpage
Does perceptual similarity predict the likelihood of a memory for the cover?

GLMM:
```{r}
summary(glmer(memory_cover ~ mean_sim + (1 | internal_id) + (1 | song_id), 
              data, family = binomial))
```


\newpage
Compute likelihood per song.
```{r}
data_by_song <- data %>%
  group_by(song_id, memory_cover) %>%
  summarize(count = n()) %>%
  pivot_wider(id_cols = song_id, names_from = memory_cover, values_from = count) %>%
  mutate(likelihood = `TRUE` / (`TRUE` + `FALSE`)) %>%
  left_join(., by_song, by = join_by(song_id))
```

```{r}
summary(lm(likelihood ~ mean_sim, data_by_song))
```




\newpage
# Features to memory likelihood

Tempo
```{r}
summary(glmer(memory_cover ~ tempo_difference + (1 | internal_id) + (1 | song_id), 
              data, family = binomial))
```

Timbre
```{r}
summary(glmer(memory_cover ~ mfccs_dist + (1 | internal_id) + (1 | song_id), 
              data, family = binomial))
```

Key
```{r}
summary(glmer(memory_cover ~ key_difference + (1 | internal_id) + (1 | song_id), 
              data, family = binomial))
```
Voice type
```{r}
summary(glmer(memory_cover ~ voice_type_difference + (1 | internal_id) + (1 | song_id), 
              data, family = binomial))
```

Valence
```{r}
summary(glmer(memory_cover ~ mean_valence_difference + (1 | internal_id) + (1 | song_id), 
              data, family = binomial))
```

Arousal
```{r}
summary(glmer(memory_cover ~ mean_arousal_difference + (1 | internal_id) + (1 | song_id), 
              data, family = binomial))
```

\newpage
# Features to memory likelihood (by song)

Tempo
```{r}
summary(lm(likelihood ~ tempo_difference, data_by_song))
```

Timbre
```{r}
summary(lm(likelihood ~ mfccs_dist, data_by_song))
```

```{r}
summary(lm(likelihood ~ key_difference, data_by_song))
```

```{r}
summary(lm(likelihood ~ voice_type_difference, data_by_song))
```

```{r}
summary(lm(likelihood ~ mean_valence_difference, data_by_song))
```

```{r}
summary(lm(likelihood ~ mean_arousal_difference, data_by_song))
```


\newpage
# Visualizations

## GLMMs

```{r}
# testing - manipulate to make sure plotting function is working
data %>%
  mutate(test_similarity = ifelse(memory_cover, mean_sim + 2, mean_sim)) %>%
  mutate(memory_cover.int = ifelse(memory_cover == 1, 1, 0)) %>%
  ggplot(aes(test_similarity, memory_cover.int)) +
  geom_point() +
  geom_jitter(width = 0, height = 0.1, alpha = 0.5) +
  xlab('FAKE similarity (for testing)') +
  ylab('memory evoked for cover') +
  stat_smooth(method = 'glm', color = 'black',
              method.args = list(family = binomial))
```


```{r}
data %>%
  mutate(memory_cover.int = ifelse(memory_cover == 1, 1, 0)) %>%
  ggplot(aes(mean_sim, memory_cover.int)) +
  geom_jitter(width = 0, height = 0.1, alpha = 0.5) +
  xlab('mean similarity') +
  ylab('memory evoked for cover') +
  stat_smooth(method = 'glm', color = 'black',
              method.args = list(family = binomial))
```

Tempo
```{r}
data %>%
  mutate(memory_cover.int = ifelse(memory_cover == 1, 1, 0)) %>%
  ggplot(aes(tempo_difference, memory_cover.int)) +
  geom_jitter(width = 0, height = 0.1, alpha = 0.5) +
  xlab('tempo difference') +
  ylab('memory evoked for cover') +
  stat_smooth(method = 'glm', color = 'black',
              method.args = list(family = binomial))
```

Voice type
```{r}
data %>%
  mutate(memory_cover.int = ifelse(memory_cover == 1, 1, 0)) %>%
  ggplot(aes(voice_type_difference, memory_cover.int)) +
  geom_jitter(width = 0.1, height = 0.1, alpha = 0.5) +
  xlab('voice type difference') +
  ylab('memory evoked for cover') +
  stat_smooth(method = 'glm', color = 'black',
              method.args = list(family = binomial))
```

## OLMs
```{r}
ggplot(data_by_song, aes(mean_sim, likelihood)) +
  geom_point() +
  geom_smooth(method = "lm", color = 'black') +
  theme_light() +
  xlab('mean similarity per song') +
  ylab('likelihood of cover memory')
#ggsave('../figures/similarity_likelihood.png', width = 7, height = 5)
```

```{r}
ggplot(data_by_song, aes(tempo_difference, likelihood)) +
  geom_point() +
  geom_smooth(method = "lm", color = 'black') +
  theme_light() +
  xlab('tempo difference per song') +
  ylab('likelihood of cover memory')
#ggsave('../figures/tempo_likelihood.png', width = 7, height = 5)
```

