---
title: "Memory content"
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

This notebook looks at memory content for pairs of original-cover memories, as predicted by perceptual similarity of the clips evoking the memories.

```{r}
library(tidyverse)
library(magrittr)
library(lme4)
library(lmerTest)
```

Load the data.
```{r}
data <- read_csv('../../data/processed/dense/joined_with_features/memory_pairs.csv')
```

Load just the features and similarity for song-wise analysis.
```{r}
by_song <- read_csv('../../data/processed/dense/joined_with_features/similarity_and_scaled_feature_differences.csv')
```

Make same/different coding a factor. (Filter out private, lump similar in with same.)
```{r}
data_factor <- data %>%
  filter(content != 'private') %>%
  mutate(content = factor(content, 
                          levels = c('same', 'similar', 'diff'), 
                          labels = c('same', 'same', 'different')))

```


Look at same/different coding vs semantic distance from sentence embeddings.

```{r}
ggplot(data, aes(x = content, y = semantic_distance)) +
  geom_jitter(width = 0.25)
```

```{r}
ggplot(data_factor, aes(x = content, y = semantic_distance)) +
  geom_jitter(width = 0.25)
```

Predict semantic distance from perceptual similarity.
```{r}
summary(lmer(semantic_distance ~ mean_sim + (1 | internal_id) + (1 | song_id),
             data_factor))
```

Predict content (same/different) from perceptual similarity.
```{r}
summary(glmer(content ~ mean_sim + (1 | internal_id) + (1 | song_id),
              data_factor, family = binomial))
```

Visualize.
```{r}
ggplot(data_factor, aes(x = mean_sim, y = semantic_distance, color = content)) +
  geom_point(alpha = 0.8)
```


Can we predict semantic distance just for same memories? (expect no)
```{r}
data_factor %>%
  filter(content == 'same') %>%
  lmer(semantic_distance ~ mean_sim + (1 | internal_id) + (1 | song_id), .) %>%
  summary()
```

Can we predict semantic distance just for different memories? (expect yes??)

```{r}
data_factor %>%
  filter(content == 'different') %>%
  lmer(semantic_distance ~ mean_sim + (1 | internal_id) + (1 | song_id), .) %>%
  summary()
```


Compute proportion of same or similar memories per song.
```{r}
data_by_song <- data_factor %>%
  group_by(song_id, content) %>%
  summarize(count = n()) %>%
  pivot_wider(id_cols = song_id, names_from = content, values_from = count) %>%
  mutate(different = ifelse(is.na(different), 0, different)) %>%
  mutate(prop_same = same / (same + different))
```

```{r}
data_by_song %<>%
  left_join(., by_song, by = join_by(song_id))
```

```{r}
summary(lm(prop_same ~ mean_sim, data_by_song))
```

Visualize.
```{r}
ggplot(data_by_song, aes(x = mean_sim, y = prop_same)) +
  geom_point() +
  geom_smooth(method = "lm", color = 'black')
```