---
title: "Memory qualities"
author: "R. Cassano-Coleman"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

This notebook looks at memory qualities (vividness, how spontaneous, etc) for pairs of original-cover memories, as predicted by perceptual similarity of the clips evoking the memories.

```{r}
library(tidyverse)
library(magrittr)
library(lme4)
library(lmerTest)
```

Load the data.
```{r}
data <- read_csv('../../data/processed/dense/joined_with_features/memory_pairs.csv')
```

Load just the features and similarity for song-wise analysis.
```{r}
by_song <- read_csv('../../data/processed/dense/joined_with_features/similarity_and_scaled_feature_differences.csv')
```

\newpage
Visualize.

```{r}
ggplot(data, aes(x = mean_sim, y = control_diff)) +
  geom_point(alpha = 0.5)
```

```{r}
ggplot(data, aes(x = mean_sim, y = vivid_diff)) +
  geom_point(alpha = 0.5)
```

```{r}
ggplot(data, aes(x = mean_sim, y = abs(control_diff))) +
  geom_point(alpha = 0.5)
```

```{r}
ggplot(data, aes(x = mean_sim, y = abs(vivid_diff))) +
  geom_point(alpha = 0.5)
```

```{r}
summary(lmer(control_diff ~ mean_sim + (1 | internal_id) + (1 | song_id),
             data))
```

```{r}
summary(lmer(vivid_diff ~ mean_sim + (1 | internal_id) + (1 | song_id),
             data))
```

Compute average vividness and degree of control difference for each song.
```{r}
data_by_song_vividness <- data %>%
  group_by(song_id) %>%
  summarize(avg_vivid_diff = mean(vivid_diff),
            count = n()) 

data_by_song_control <- data %>%
  group_by(song_id) %>%
  summarize(avg_control_diff = mean(control_diff),
            count = n())  
```


```{r}
data_by_song_vividness %<>%
  left_join(., by_song, by = join_by(song_id))

data_by_song_control %<>%
  left_join(., by_song, by = join_by(song_id))
```

\newpage
```{r}
summary(lm(avg_vivid_diff ~ mean_sim, data_by_song_vividness))
```

Visualize.
```{r}
ggplot(data_by_song_vividness, aes(x = mean_sim, y = avg_vivid_diff)) +
  geom_point() +
  #geom_point(aes(size = count), alpha = 0.75) +
  geom_smooth(method = "lm", color = 'black') +
  geom_hline(aes(yintercept = 0)) +
  theme_light() +
  xlab('mean similarity per song') +
  ylab('average difference in memory vividness')
#ggsave('../figures/similarity_vividness.png', width = 7, height = 5)
```

\newpage
```{r}
summary(lm(avg_control_diff ~ mean_sim, data_by_song_control))
```

Visualize.
```{r}
ggplot(data_by_song_control, aes(x = mean_sim, y = avg_control_diff)) +
  #geom_point(aes(size = count), alpha = 0.75) +
  geom_point() +
  geom_smooth(method = "lm", color = 'black') +
  geom_hline(aes(yintercept = 0)) +
  theme_light() +
  xlab('mean similarity per song') +
  ylab('average difference in memory degree of control per song\n(how spontaneously it was retrieved)')
#ggsave('../figures/similarity_control.png', width = 7, height = 5)
```